%%\subsection{Intercepting proxy solutions and reverse proxies}

Within enterprise networks and corporations with increased levels of paranoia or at least some defined security requirements it is common \textbf{not} to allow direct connections to the public internet.

For this reason proxy solutions are deployed on corporate networks to intercept and scan the traffic for potential threats within sessions.\\

For encrypted traffic there are four options:

\begin{itemize}
\item Block the connection because it cannot be scanned for threats.
\item Bypass the threat-mitigation and pass the encrypted session to the client, which results in a situation where malicious content is transferred directly to the client without visibility to the security system.
\item Intercept (i.e. terminate) the session at the proxy, scan there and re-encrypt the session towards the client (effectively MITM).
\item Deploy special Certificate Authorities to enable Deep Packet Inspection on the wire.
\end{itemize}

While the latest solution might be the most "up to date", it arises a new front in the context of this paper, because the most secure part of a client's connection could only be within the corporate network, if the proxy-server handles the connection to the destination server in an insecure manner.

Conclusion: Don't forget to check your proxy solutions SSL-capabilities. Also do so for your reverse proxies!

\subsubsection{squid}


As of squid-3.2.7 (01 Feb 2013) there is support for the OpenSSL NO\_Compression option within squid config (CRIME attack) and if you combine that in the config file, with an enforcement of the server cipher preferences (BEAST Attack) you are safe.

\paragraph*{squid.conf}\mbox{}\\
squid.conf

\todo{UNTESTED!}
\begin{lstlisting}
options=NO_SSLv2,NO_TLSv1,NO_Compression,CIPHER_SERVER_PREFERENCE 
cipher=%*\cipherstringB*)
\end{lstlisting}



\paragraph*{squid.conf}\mbox{}\\
squid.conf

%% http://forum.pfsense.org/index.php?topic=63262.0

\todo{UNTESTED!}
\begin{lstlisting}
		NO_SSLv2    Disallow the use of SSLv2
		NO_SSLv3    Disallow the use of SSLv3
		NO_TLSv1    Disallow the use of TLSv1.0
		NO_TLSv1_1  Disallow the use of TLSv1.1
		NO_TLSv1_2  Disallow the use of TLSv1.2
		SINGLE_DH_USE
				Always create a new key when using temporary/ephemeral
				DH key exchanges
\end{lstlisting}

\todo{Patch here? Definitely working for 3.2.6!}
For squid Versions before 3.2.7 use this patch against a vanilla source-tree:

\begin{lstlisting}
--- support.cc.ini      2013-01-09 02:41:51.000000000 +0100
+++ support.cc  2013-01-21 16:13:32.549383848 +0100
@@ -400,6 +400,11 @@
         "NO_TLSv1_2", SSL_OP_NO_TLSv1_2
     },
 #endif
+#ifdef SSL_OP_NO_COMPRESSION
+    {
+        "NO_Compression", SSL_OP_NO_COMPRESSION
+    },
+#endif
     {
         "", 0
     },
\end{lstlisting}



\subsubsection{Bluecoat}
%% https://kb.bluecoat.com/index?page=content&id=KB5549
\begin{description}
\item[Tested with Version:] SGOS 6.5.x

BlueCoat Proxy SG Appliances can be used as forward and reverse proxies. The reverse proxy feature is rather under-developed, and while it is possible and supported, there only seems to be limited use of this feature "in the wild" - nonetheless there are a few cipher suites to choose from, when enabling SSL features.

\item[Only allow TLS 1.0,1.1 and 1.2 protocols:] \mbox{}

\begin{lstlisting}
$conf t
$(config)ssl
$(config ssl)edit ssl-device-profile default
$(config device-profile default)protocol tlsv1 tlsv1.1 tlsv1.2
  ok
\end{lstlisting}

\item[Select your accepted cipher-suites:] \mbox{}

\begin{lstlisting}
$conf t
Enter configuration commands, one per line.  End with CTRL-Z.
$(config)proxy-services
$(config proxy-services)edit ReverseProxyHighCipher
$(config ReverseProxyHighCipher)attribute cipher-suite
Cipher#  Use        Description        Strength
-------  ---  -----------------------  --------
      1  yes            AES128-SHA256      High
      2  yes            AES256-SHA256      High
      3  yes               AES128-SHA    Medium
      4  yes               AES256-SHA      High
      5  yes       DHE-RSA-AES128-SHA      High
      6  yes       DHE-RSA-AES256-SHA      High
               [...]
     13  yes          EXP-RC2-CBC-MD5    Export

Select cipher numbers to use, separated by commas: 2,5,6
  ok
\end{lstlisting}

The same protocols are available for forward proxy settings and should be adjusted accordingly:
In your local policy file add the following section:
\begin{lstlisting}
<ssl>
    DENY server.connection.negotiated_ssl_version=(SSLV2, SSLV3)
\end{lstlisting}

Disabling protocols and ciphers in a forward proxy environment could lead to unexpected results on certain (misconfigured?) webservers (i.e. ones accepting only SSLv2/3 protocol connections)

\end{description} 




\subsubsection{Pound}
% See http://www.apsis.ch/pound
% See https://help.ubuntu.com/community/Pound

Pound 2.6
	
\begin{lstlisting}
# HTTP Listener, redirects to HTTPS
ListenHTTP
    Address 10.10.0.10
    Port    80
    Service
        Redirect "https://some.site.tld
    End
End
## HTTPS Listener
ListenHTTPS
    Address      10.10.0.10
    Port         443
    AddHeader    "Front-End-Https: on"
    Cert         "/path/to/your/cert.pem"
    ## See 'man ciphers'.
    Ciphers      "TLSv1.2:TLSv1.1:!SSLv3:!SSLv2:%*\cipherstringB*)"
    Service
        BackEnd
            Address 10.20.0.10
            Port 80
        End
    End
End
\end{lstlisting}
