%%\subsection{Choosing your own cipher suites}
\label{section:ChoosingYourOwnCipherSuites}

\todo{ Adi...  you want to describe how to make your own selection of cipher suites here.}

SSL/TLS cipher suites consist of a key exchange mechanism, an authentication, a
stream cipher (or a block cipher with a chaining mode) and a message authentication
mechanism.

Many of those mechanisms are interchangeable like the key exchange in this example:
\texttt{ECDHE-RSA-AES256-GCM-SHA384} and \texttt{DHE-RSA-AES256-GCM-SHA384}.
To provide a decent level of security, all algorithms need to be safe (subject to
the disclaimer in section \ref{section:disclaimer}).

Note: There are some very weak cipher suites in about every crypto library, most of
them for historic reasons like the crypto export embargo
\footnote{\url{http://en.wikipedia.org/wiki/Export_of_cryptography_in_the_United_States}}.
For the following chapter support of those is assumed to be disabled by having
\texttt{!EXP:!LOW:!NULL} as part of the cipher string.

\todo{Team: do we need references for all cipher suites considered weak?}

\subsubsection{key exchange}

Many algorithms allow a secure key exchange. Among those are RSA, DSA, DH, EDH, ECDSA,
ECDH, EECDH and a few others. During the key exchange, keys for authentication and for
encryption are exchanged. For RSA and DSA those keys are the same.

\todo{explain this section}

\begin{center}
\begin{tabular}{| l | l | l | l |}
    \toprule
 & \textbf{Key}  & \textbf{\cellcolor{orange}EC}  & \textbf{\cellcolor{green}ephemeral} \\ \cmidrule(lr){1-4}
    \cellcolor{red}    RSA   & RSA  & \cellcolor{green}no   & \cellcolor{red} no         \\
    \cellcolor{red}    DH    & RSA  & \cellcolor{green}no   & \cellcolor{red} no         \\
    \cellcolor{green}  EDH   & RSA  & \cellcolor{green}no   & \cellcolor{green} yes      \\
    \cellcolor{red}    ECDH  & both & \cellcolor{orange}yes & \cellcolor{red} no         \\
    \cellcolor{orange} EECDH & both & \cellcolor{orange}yes & \cellcolor{green} yes      \\
    \cellcolor{red}    DSA   & DSA  & \cellcolor{green}no   & \cellcolor{red} no         \\
    \cellcolor{red}    ECDSA & DSA  & \cellcolor{orange}yes & \cellcolor{red} no         \\
\bottomrule
\end{tabular}
%\\
%\\
%disabled: \texttt{!PSK:!SRP}
\end{center}

\textbf{Ephemeral Key Exchange} uses different keys for authentication (the server's RSA
key) and encryption (a randomly created key). This advantage is called ``Forward
Secrecy'' and means that even recorded traffic cannot be decrypted later when someone
gets the server key. \\
All ephemeral key exchange mechanisms base on Diffie-Hellman algorithm and require
pre-generated Diffe-Hellman parameter (which allow fast ephemeral key generation). It
is important to note that the Diffie-Hellman parameters need to be at least as strong
(speaking in number of bits) as the RSA host key. \todo{TODO: reference!}


\textbf{Elliptic Curves}\ref{section:EllipticCurveCryptography} required by current TLS
standards only consist of the so-called NIST-curves (\texttt{secp256r1} and
\texttt{secp384r1}) which may be weak because the parameters that led to their generation
weren't properly explained (by the NSA). \\
Disabling support for Elliptic Curves leads to no ephemeral key exchange being available
for the Windows platform. When you decide to use Elliptic Curves despite the uncertainty,
make sure to at least use the stronger curve of the two supported by all clients
(\texttt{secp384r1}).


Other key exchange mechanisms like Pre-Shared Key (PSK) or Secure Remote Password
(SRP) are irrelevant for regular SSL/TLS use.

\subsubsection{authentication}

RSA, DSA, DSS, ECDSA, ECDH, FORTEZZA(?).

Other authentication mechanisms like Pre Shared Keys aren't used in SSL/TLS: \texttt{!PSK:!aNULL}

\subsubsection{encryption}

AES, CAMELLIA, SEED, ARIA(?), FORTEZZA(?)...

Other ciphers like IDEA, RC2, RC4, 3DES or DES are weak and therefor not recommended:
\texttt{!DES:!3DES:!RC2:!RC4:!eNULL}

\subsubsection{message authentication}

SHA-1 (SHA), SHA-2 (SHA256, SHA384), AEAD

Note that SHA-1 is considered broken and should not be used. SHA-1 is however a the
only still available message authentication mechanism supporting TLS1.0/SSLv3. Without
SHA-1 most clients will be locked out.

Other hash functions like MD2, MD4 or MD5 are unsafe and broken: \texttt{!MD2:!MD4:!MD5}

\subsubsection{combining cipher strings}
%% reference 'man ciphers' and 'openssl ciphers' and show some simple examples
%% VERY IMPORTANT: hint at the IANA-list and the differences in implementations

\todo{ Adi...  The text below was simply the old text, still left here for reference.}

%%% NOTE: we do not need to list this all here, can move to an appendix
%At the time of this writing, SSL is defined in RFCs: 	
%
%\begin{itemize}
%\item RFC2246 - TLS1.0		
%\item RFC3268 - AES		
%\item RFC4132 - Camelia		
%\item RFC4162 - SEED		
%\item RFC4279 - PSK		
%\item RFC4346 - TLS 1.1		
%\item RFC4492 - ECC		
%\item RFC4785 - PSK\_NULL		
%\item RFC5246 - TLS 1.2		
%\item RFC5288 - AES\_GCM		
%\item RFC5289 - AES\_GCM\_SHA2\_ECC		
%\item RFC5430 - Suite B		
%\item RFC5487 - GCM\_PSK		
%\item RFC5489 - ECDHE\_PSK		
%\item RFC5932 - Camelia		
%\item RFC6101 - SSL 3.0		
%\item RFC6209 - ARIA		
%\item RFC6367 - Camelia		
%\item RFC6655 - AES\_CCM		
%\item RFC7027 - Brainpool Curves		
%\end{itemize}

\subsubsection{Overview of SSL Server settings}


Most Server software (Webservers, Mail servers, etc.) can be configured to prefer certain cipher suites over others. 
We followed the recommendations by Ivan Ristic's SSL/TLS Deployment Best Practices\footnote{\url{https://www.ssllabs.com/projects/best-practices/index.html}} document (see section 2.2 "Use Secure Protocols") and arrived at a list of recommended cipher suites for SSL enabled servers.

Following Ivan Ristic's adivce we arrived at a categorisation of cipher suites.

\begin{center}
\begin{tabular}{lllll}
\cmidrule[\heavyrulewidth]{2-5}
& \textbf{Version}   & \textbf{KeyEx} & \textbf{Cipher}    & \textbf{MAC}       \\\cmidrule(lr){2-5}
\cellcolor{green}prefer  & TLS 1.2   & DHE\_DSS   & AES\_256\_GCM   & SHA384        \\
    &   & DHE\_RSA   & AES\_256\_CCM   & SHA256        \\
    &   & ECDHE\_ECDSA   & AES\_256\_CBC   &       \\
    &   & ECDHE\_RSA &   &       \\ 
    &   &   &   &       \\
\cellcolor{orange}consider    & TLS 1.1   & DH\_DSS    & AES\_128\_GCM   & SHA       \\
    & TLS 1.0   & DH\_RSA    & AES\_128\_CCM   &       \\
    &   & ECDH\_ECDSA    & AES\_128\_CBC   &       \\ 
    &   & ECDH\_RSA  & CAMELLIA\_256\_CBC  &       \\
    &   & RSA   & CAMELLIA\_128\_CBC  &       \\
    &   &   &   &       \\
\cellcolor{red}avoid   
& SSL 3.0   & NULL  & NULL  & NULL      \\
    &   & DH\_anon   & RC4\_128   & MD5       \\
    &   & ECDH\_anon & 3DES\_EDE\_CBC  &       \\
    &   &   & DES\_CBC   &       \\
    &   &   &   &       \\
\cellcolor{blue}{\color{white}special }
&   & PSK   & CAMELLIA\_256\_GCM  &       \\
    &   & DHE\_PSK   & CAMELLIA\_128\_GCM  &       \\
    &   & RSA\_PSK   & ARIA\_256\_GCM  &       \\
    &   & ECDHE\_PSK & ARIA\_256\_CBC  &       \\
    &   &   & ARIA\_128\_GCM  &       \\
    &   &   & ARIA\_128\_CBC  &       \\
    &   &   & SEED  &       \\
\cmidrule[\heavyrulewidth]{2-5}
\end{tabular}
\end{center}

A remark on the ``consider'' section: the BSI (Federal office for information security, Germany) recommends in its technical report TR-02102-2\footnote{\url{https://www.bsi.bund.de/SharedDocs/Downloads/DE/BSI/Publikationen/TechnischeRichtlinien/TR02102/BSI-TR-02102-2_pdf.html}} to \textbf{avoid} non-ephemeral\footnote{Ephemeral keys are session keys which are destroyed upon termination of the encrypted session. In TLS/SSL, they are realized by the DHE cipher suites. } keys for any communication which might contain personal or sensitive data. In this document, we follow BSI's advice and therefore only keep cipher suites containing (EC)DH\textbf{E} (ephemeral) variants. System administrators, who can not use forward secrecy can still use the cipher suites in the ``consider'' section. We however, do not recommend them in this document.

%% NOTE: s/forward secrecy/perfect forward secrecy???

Note that the entries marked as ``special'' are cipher suites which are not common to all clients (webbrowsers etc).


\subsubsection{Tested clients}
 
Next we tested the cipher suites above on the following clients:

%% NOTE: we need to test with more systems!!
\begin{itemize}
\item Chrome 30.0.1599.101 Mac OS X 10.9
\item Safari 7.0 Mac OS X 10.9
\item Firefox 25.0 Mac OS X 10.9
\item Internet Explorer 10 Windows 7
\item Apple iOS 7.0.3
\end{itemize}


The result of testing the cipher suites with these clients gives us a preference order as shown in table \ref{table:prefOrderCipherSuites}. 
Should a client not be able to use a specific cipher suite, it will fall back to the next possible entry as given by the ordering.

\begin{table}[h]
\centering\small
    \begin{tabular}{cllcccc}
    \toprule
    \textbf{Pref}   & \textbf{Cipher Suite}                            & \textbf{ID}   & \multicolumn{4}{l}{\textbf{Supported by}}\\ 
    \cmidrule(lr){4-7}
                    & \textbf{OpenSSL Name}                            &               & Chrome & FF   & IE   & Safari \\
    \cmidrule(lr){1-7}
    \phantom{0}1    & \verb|TLS_DHE_RSA_WITH_AES_256_GCM_SHA384|     & \verb|0x009f| & \no    & \no  & \no  & \no    \\
                    & \verb|DHE-RSA-AES256-GCM-SHA384|                      &               & &&&\\\rowcolor{lightlightgray}
    \phantom{0}2    & \verb|TLS_ECDHE_ECDSA_WITH_AES_256_CBC_SHA384| & \verb|0xC024| & \no    & \no  & \no  & \yes   \\\rowcolor{lightlightgray}
                    & \verb|ECDHE-ECDSA-AES256-SHA384|                      &               & &&&\\
    \phantom{0}3    & \verb|TLS_ECDHE_RSA_WITH_AES_256_CBC_SHA384|   & \verb|0xC028| & \no    & \no  & \no  & \yes   \\
                    & \verb|ECDHE-RSA-AES256-SHA384|                        &               & &&&\\\rowcolor{lightlightgray}
    \phantom{0}4    & \verb|TLS_DHE_RSA_WITH_AES_256_CBC_SHA256|     & \verb|0x006B| & \yes   & \no  & \no  & \yes   \\\rowcolor{lightlightgray}
                    & \verb|DHE-RSA-AES256-SHA256|                          &               & &&&\\
    \phantom{0}5    & \verb|TLS_ECDHE_ECDSA_WITH_AES_256_CBC_SHA|    & \verb|0xC00A| & \yes   & \yes & \yes & \yes   \\
                    & \verb|ECDHE-ECDSA-AES256-SHA|                         &               & &&&\\\rowcolor{lightlightgray}
    \phantom{0}6    & \verb|TLS_ECDHE_RSA_WITH_AES_256_CBC_SHA|      & \verb|0xC014| & \yes   & \yes & \yes & \yes   \\\rowcolor{lightlightgray}
                    & \verb|ECDHE-RSA-AES256-SHA|                           &               & &&&\\
    \phantom{0}7    & \verb|TLS_DHE_RSA_WITH_AES_256_CBC_SHA|        & \verb|0x0039| & \yes   & \yes & \no  & \yes   \\
                    & \verb|DHE-RSA-AES256-SHA|                             &               & &&&\\\rowcolor{lightlightgray}
    \phantom{0}8    & \verb|TLS_DHE_DSS_WITH_AES_256_CBC_SHA|        & \verb|0x0038| & \no    & \yes & \yes & \no    \\\rowcolor{lightlightgray}
                    & \verb|DHE-DSS-AES256-SHA|                             &               & &&&\\
    \phantom{0}9    & \verb|TLS_DHE_RSA_WITH_CAMELLIA_256_CBC_SHA|   & \verb|0x0088| & \no    & \yes & \no  & \no    \\
                    & \verb|DHE-RSA-CAMELLIA256-SHA|                        &               & &&&\\\rowcolor{lightlightgray}
    \phantom{}10    & \verb|TLS_DHE_DSS_WITH_CAMELLIA_256_CBC_SHA|   & \verb|0x0087| & \no    & \yes & \no  & \no    \\\rowcolor{lightlightgray}
                    & \verb|DHE-DSS-CAMELLIA256-SHA|                        &               & &&&\\
   \bottomrule
    \end{tabular}
\caption{Preference order of cipher suites.  All suites are supported by OpenSSL.}
\label{table:prefOrderCipherSuites}
\end{table}

Note: the above table \ref{table:prefOrderCipherSuites} contains Elliptic curve key exchanges. There are currently strong doubts\footnote{\url{http://safecurves.cr.yp.to/rigid.html}} concerning ECC.
If unsure, remove the cipher suites starting with ECDHE in the table above.


Based on this ordering, we can now define the corresponding settings for servers. We will start with the most common web servers.
