\documentclass{nlctdoc}

\usepackage[utf8]{inputenc}
\ifpdf
\usepackage[T1]{fontenc}
\usepackage{mathpazo}
\usepackage[scaled=.88]{helvet}
\usepackage{courier}
\usepackage{xr-hyper}
\fi

\usepackage{alltt}
\usepackage{pifont}
\usepackage{array}

\usepackage[colorlinks,pdfauthor={Nicola L.C. Talbot},
            pdftitle={The glossaries package: a guide for beginners}]{hyperref}
\usepackage[nomain]{glossaries}

\newcommand*{\yes}{\ding{52}}
\newcommand*{\no}{\ding{56}}

\title{The glossaries package: a guide for beginners}
\author{Nicola L.C. Talbot}
\date{2014-04-04}

\ifpdf
  \externaldocument{glossaries-user}
\fi

\makeatletter
\newcommand*{\optionlabel}[1]{%
 \@glstarget{option#1}{}\textbf{Option~#1}}
\makeatother

\newcommand*{\opt}[1]{\hyperlink{option#1}{Option~#1}}
\newcommand*{\optsor}[2]{Options~\hyperlink{option#1}{#1}
or~\hyperlink{option#2}{#2}}
\newcommand*{\optsand}[2]{Options~\hyperlink{option#1}{#1}
and~\hyperlink{option#2}{#2}}

\begin{document}
\maketitle

\begin{abstract}
This document is a brief guide to the \styfmt{glossaries} package
for beginners who find the size of the main user manual daunting
and, as such, it only covers the basics. For brevity some options to
the commands described here are omitted. For a~more detailed guide,
see the \docref{main user manual}{glossaries-user}.
\end{abstract}

\tableofcontents

\section{Defining Terms}
\label{sec:defterm}

When you use the \sty{glossaries} package, you need to define
glossary entries in the document preamble. These entries could be
a~word, phrase, acronym or symbol. They're usually accompanied by
a~description, which could be a short sentence or an in-depth
explanation that spans multiple paragraphs. The simplest method of
defining an entry is to use:
\begin{definition}
\begin{alltt}
\cs{newglossaryentry}\marg{label}
\verb|{|
  name=\marg{name},
  description=\marg{description},
  \meta{other options}
\verb|}|
\end{alltt}
\end{definition}
where \meta{label} is a unique label that identifies this entry.
(Don't include the angle brackets \meta{ }. They just indicate the parts of
the code you need to change when you use this command in your document.) As
with similar labelling commands, such as \cs{label} or \cs{bibitem},
the label should not contain active characters, so just use
\texttt{a}, \ldots, \texttt{z}, \texttt{A}, \ldots, \texttt{Z},
\texttt{0}, \ldots, \texttt{9}. You may also be able to use some
punctuation characters, unless they have been made active (for
example, via \sty{babel}'s shorthand activation.)
The \meta{name} is the word, phrase or symbol you are defining,
and \meta{description} is the description.

This command is a ``short'' command, which means that
\meta{description} can't contain a~paragraph break. If you have
a~long description, you can instead use:
\begin{definition}
\begin{alltt}
\cs{longnewglossaryentry}\marg{label}
\verb|{|
  name=\marg{name},
  \meta{other options}
\verb|}|
\marg{description}
\end{alltt}
\end{definition}

Examples:
\begin{enumerate}
\item Define the term ``set'' with the label \texttt{set}:
\begin{verbatim}
\newglossaryentry{set}
{
  name={set},
  description={a collection of objects}
}
\end{verbatim}

\item Define the symbol $\emptyset$ with the label
\texttt{emptyset}:
\begin{verbatim}
\newglossaryentry{emptyset}
{
  name={\ensuremath{\emptyset}},
  description={the empty set}
}
\end{verbatim}

\item Define the phrase ``Fish Age'' with the label
\texttt{fishage}:
\begin{verbatim}
\longnewglossaryentry{fishage}
{name={Fish Age}}
{%
  A common name for the Devonian geologic period 
  spanning from the end of the Silurian Period to
  the beginning of the Carboniferous Period.

  This age was known for its remarkable variety of 
  fish species.
}
\end{verbatim}
(The percent character discards the end of line character that would
otherwise cause an unwanted space to appear at the start of the
description.)

\item Take care if the first letter is an extended Latin or
non-Latin character (either specified via a~command such as 
\verb|\'e| or explicitly via the \sty{inputenc} package such 
as \texttt{\'e}). This first letter must be placed in a~group:

\begin{verbatim}
\newglossaryentry{elite}
{
  name={{\'e}lite},
  description={select group or class}
}
\end{verbatim}
or
\begin{alltt}
\verb|\newglossaryentry{elite}|
\verb|{|
  name=\verb|{{|\'e\verb|}lite}|,
  description=\verb|{select group or class}|
\verb|}|
\end{alltt}
\end{enumerate}

Acronyms or abbreviations can be defined using
\begin{definition}
\cs{newacronym}\marg{label}\marg{short}\marg{long}
\end{definition}
where \meta{label} is the label (as with the \cs{newglossaryentry}
and the \cs{longnewglossaryentry} commands), \meta{short} is the abbreviation or
acronym and \meta{long} is the long form. For example:
\begin{verbatim}
\newacronym{svm}{svm}{support vector machine}
\end{verbatim}
This defines a glossary entry with the label \texttt{svm}. By
default, the \meta{name} is set to \meta{short} (``svm'' in the
above example) and the \meta{description} is set to \meta{long}
(``support vector machine'' in the above example). If, instead, you
want to be able to specify your own description you can do this
using the optional argument:
\begin{verbatim}
\newacronym
 [description={statistical pattern recognition technique}]
 {svm}{svm}{support vector machine}
\end{verbatim}

There are other keys you can use when you define an entry. For
example, the \texttt{name} key used above indicates how the term
should appear in the list of entries (glossary). If the term should
appear differently when you reference it in the document, you need
to use the \texttt{text} key as well.

For example:
\begin{verbatim}
\newglossaryentry{latinalph}
{
  name={Latin Alphabet},
  text={Latin alphabet},
  description={alphabet consisting of the letters 
  a, \ldots, z, A, \ldots, Z}
}
\end{verbatim}
This will appear in the text as \qt{Latin alphabet} but will be listed in
the glossary as \qt{Latin Alphabet}.

Another commonly used key is \texttt{plural} for specifying the
plural of the term. This defaults to the value of the \texttt{text}
key with an ``s'' appended, but if this is incorrect, just use the
\texttt{plural} key to override it:
\begin{verbatim}
\newglossaryentry{oesophagus}
{
  name={{\oe}sophagus},
  plural={{\oe}sophagi},
  description={canal from mouth to stomach}
}
\end{verbatim}
(Remember from earlier that the initial ligature \cs{oe} needs to
be grouped.)

The plural forms for acronyms can be specified using the
\texttt{longplural} and \texttt{shortplural} keys. For example:
\begin{verbatim}
\newacronym
 [longplural={diagonal matrices}]
 {dm}{DM}{diagonal matrix}
\end{verbatim}
If omitted, the defaults are again obtained by appending an ``s'' to
the singular versions.

It's also possible to have both a~name and a~corresponding symbol.
Just use the \texttt{name} key for the name and the \texttt{symbol} key
for the symbol. For example:
\begin{verbatim}
\newglossaryentry{emptyset}
{
  name={empty set},
  symbol={\ensuremath{\emptyset}},
  description={the set containing no elements}
}
\end{verbatim}

\section{Using Entries}
\label{sec:useterm}

Once you have defined your entries, as described above, you can
reference them in your document. There are a~number of commands to
do this, but the most common one is:
\begin{definition}
\cs{gls}\marg{label}
\end{definition}
where \meta{label} is the label you assigned to the entry when you
defined it. For example, \verb|\gls{fishage}| will display \qt{Fish
Age} in the text (given the definition from the previous section).

If the entry was defined as an acronym (using \cs{newacronym}
described above), then \cs{gls} will display the full form the first
time it's used and just the short form on
subsequent use. For example, \verb|\gls{svm}| will display \qt{support vector
machine (svm)} the first time it's used, but the next occurrence of
\verb|\gls{svm}| will just display \qt{svm}.

If you want the plural form, you can use:
\begin{definition}
\cs{glspl}\marg{label}
\end{definition}
instead of \cs{gls}\marg{label}. For example, \verb|\glspl{set}|
displays \qt{sets}.

If the term appears at the start of a~sentence, you can convert the
first letter to upper case using:
\begin{definition}
\cs{Gls}\marg{label}
\end{definition}
for the singular form or
\begin{definition}
\cs{Glspl}\marg{label}
\end{definition}
for the plural form. For example:
\begin{verbatim}
\Glspl{set} are collections.
\end{verbatim}
produces \qt{Sets are collections}.

If you've specified a symbol using the \texttt{symbol} key, you can
display it using:
\begin{definition}
\cs{glssymbol}\marg{label}
\end{definition}

\section{Acronyms}
\label{sec:acronyms}

Recall from above, the first time you use an acronym
with \cs{gls}, it's full form is displayed but subsequent uses
display only the short form. By default, the first use displays
\meta{long} (\meta{short}). That is, the long form is displayed
followed by the short form in parentheses. You can change this
first-use format with:
\begin{definition}
\cs{setacronymstyle}\marg{style name}
\end{definition}
(This must be used before you start defining your acronyms with
\cs{newacronym}.) There are a~number of predefined styles listed in
\xrsectionref{sec:predefinedacrstyles}{glossaries-user}{Predefined
Acronym Styles} in the main \sty{glossaries} user manual. Here are
a~few examples:
\begin{enumerate}
\item \meta{short} (\meta{long})
\begin{verbatim}
\setacronymstyle{short-long}
\end{verbatim}
This displays the short form followed by the long form in
parentheses.

\item \meta{long} (\cs{textsc}\marg{short})
\begin{verbatim}
\setacronymstyle{long-short-sc}
\end{verbatim}
This is like the default style but the short form is displayed in
small caps. Remember that when you use \cs{textsc}\marg{text} to generate small
capitals, you must specify \meta{text} in lower case, so the short
form of the acronym should be defined in lower case. For example:
\begin{verbatim}
\newacronym{svm}{svm}{support vector machine}
\end{verbatim}

\item \meta{long} (\cs{textsmaller}\marg{short})
\begin{verbatim}
\setacronymstyle{long-short-sm}
\end{verbatim}
This is similar to the previous style but uses \cs{textsmaller} to
format the short form. Remember to load the \sty{relsize} package,
which defines \cs{textsmaller}, if you want to use this style. The
short form now needs to be defined in upper case:
\begin{verbatim}
\newacronym{svm}{SVM}{support vector machine}
\end{verbatim}

\end{enumerate}

\section{Displaying a List of Entries}
\label{sec:printglossaries}

Suppose you now want to display a list of all the entries you've
referenced in your document. This is where things start to get
complicated and a~lot of new users get bewildered. You have three
options:

\begin{description}
\item[]\optionlabel1: 

 This is the simplest option but it's slow and if
 you want a sorted list, it doesn't work for non-Latin alphabets.

  \begin{enumerate}
    \item Add \cs{makenoidxglossaries} to your preamble (before you
    start defining your entries, as described in
    \sectionref{sec:defterm}).

    \item Put
\begin{definition}
\cs{printnoidxglossary}[sort=\meta{order},\meta{other options}]
\end{definition}
    where you want your list of entries to appear. The sort
    \meta{order} may be one of: \texttt{word} (word ordering),
    \texttt{letter} (letter ordering), \texttt{case} (case-sensitive
    letter ordering), \texttt{def} (in order of definition) or
    \texttt{use} (in order of use).

    \item Run \LaTeX\ twice on your document. (As you would do to
    make a~table of contents appear.) For example, click twice on
    the ``typeset'' or ``build'' or ``PDF\LaTeX'' button in your editor.
  \end{enumerate}

\item[]\optionlabel2:

   This option uses an application called \texttt{makeindex} to sort 
   the entries. This application comes with all modern \TeX\ distributions, 
   but it's hard-coded for the non-extended Latin alphabet. This process 
   involves making \LaTeX\ write the glossary information to a temporary 
   file which \texttt{makeindex} reads. Then \texttt{makeindex} writes 
   a~new file containing the code to typeset the glossary. \LaTeX\ then 
   reads this file on the next run.

   \begin{enumerate}
    \item Add \cs{makeglossaries} to your preamble (before you start
    defining your entries).

    \item Put
\begin{definition}
\cs{printglossary}\oarg{options}
\end{definition}
    where you want your list of entries (glossary) to appear.

    \item Run \LaTeX\ on your document. This creates files with the
    extensions \texttt{.glo} and \texttt{.ist} (for example, if your 
    \LaTeX\ document is called \texttt{myDoc.tex}, then you'll have 
    two extra files called \texttt{myDoc.glo} and \texttt{myDoc.ist}).
    If you look at your document at this point, you won't see the 
    glossary as it hasn't been created yet.

    \item Run \texttt{makeindex} with the \texttt{.glo} file as the
    input file and the \texttt{.ist} file as the style so that
    it creates an output file with the extension \texttt{.gls}. If
    you have access to a terminal or a command prompt (for example, the
    MSDOS command prompt for Windows users or the bash console for
    Unix-like users) then you need to run the command:
\begin{verbatim}
makeindex -s myDoc.ist -o myDoc.gls myDoc.glo
\end{verbatim}
   (Replace \texttt{myDoc} with the base name of your \LaTeX\
    document file. Avoid spaces in the file name.) If you don't know
    how to use the command prompt, then you can probably access
    \texttt{makeindex} via your text editor, but each editor has a
    different method of doing this, so I~can't give a~general
    description. You will have to check your editor's manual.

    The default sort is word order (``sea lion'' comes before ``seal''). 
    If you want letter ordering you need to add the \texttt{-l}
    switch:
\begin{verbatim}
makeindex -l -s myDoc.ist -o myDoc.gls myDoc.glo
\end{verbatim}

    \item Once you have successfully completed the previous step,
    you can now run \LaTeX\ on your document again.
   \end{enumerate}

\item[]\optionlabel3:

   This option uses an application called
   \texttt{xindy} to sort the entries. This application is more
   flexible than \texttt{makeindex} and is able to sort extended
   Latin or non-Latin alphabets. It comes with \TeX~Live but not
   with MiK\TeX. Since \texttt{xindy} is a Perl script, if you are
   using MiK\TeX\ you will not only need to install \texttt{xindy}, you
   will also need to install Perl. In a~similar way to \opt2, this 
   option involves making \LaTeX\ write the glossary information to 
   a~temporary file which \texttt{xindy} reads. Then \texttt{xindy} 
   writes a~new file containing the code to typeset the glossary. 
   \LaTeX\ then reads this file on the next run.

   \begin{enumerate}
     \item Add the \texttt{xindy} option to the \sty{glossaries}
package option list:
\begin{verbatim}
\usepackage[xindy]{glossaries}
\end{verbatim}

     \item Add \cs{makeglossaries} to your preamble (before you start
     defining your entries).

    \item Put
\begin{definition}
\cs{printglossary}\oarg{options}
\end{definition}
    where you want your list of entries (glossary) to appear.

    \item Run \LaTeX\ on your document. This creates files with the
    extensions \texttt{.glo} and \texttt{.xdy} (for example, if your 
    \LaTeX\ document is called \texttt{myDoc.tex}, then you'll have 
    two extra files called \texttt{myDoc.glo} and \texttt{myDoc.xdy}).
    If you look at your document at this point, you won't see the 
    glossary as it hasn't been created yet.

    \item Run \texttt{xindy} with the \texttt{.glo} file as the
    input file and the \texttt{.xdy} file as a~module so that
    it creates an output file with the extension \texttt{.gls}. You 
    also need to set the language name and input encoding. If
    you have access to a terminal or a command prompt (for example, the
    MSDOS command prompt for Windows users or the bash console for
    Unix-like users) then you need to run the command (all on one
    line):
\begin{verbatim}
xindy  -L english -C utf8 -I xindy -M myDoc 
-t myDoc.glg -o myDoc.gls myDoc.glo
\end{verbatim}
    (Replace \texttt{myDoc} with the base name of your \LaTeX\
    document file. Avoid spaces in the file name. If necessary, also replace
    \texttt{english} with the name of your language and \texttt{utf8}
    with your input encoding.) If you don't know
    how to use the command prompt, then you can probably access
    \texttt{xindy} via your text editor, but each editor has a
    different method of doing this, so I~can't give a~general
    description. You will have to check your editor's manual.

    The default sort is word order (``sea lion'' comes before ``seal''). 
    If you want letter ordering you need to add the
    \texttt{order=letter} package option:
\begin{verbatim}
\usepackage[xindy,order=letter]{glossaries}
\end{verbatim}

    \item Once you have successfully completed the previous step,
    you can now run \LaTeX\ on your document again.

   \end{enumerate}

\end{description}

For \optsand23, it can be difficult to remember all the
parameters required for \texttt{makeindex} or \texttt{xindy}, so the
\sty{glossaries} package provides a~script called
\texttt{makeglossaries} that reads the \texttt{.aux} file to
determine what settings you have used and will then run
\texttt{makeindex} or \texttt{xindy}. Again, this is a~command line
application and can be run in a~terminal or command prompt. For
example, if your \LaTeX\ document is in the file \texttt{myDoc.tex},
then run:
\begin{verbatim}
makeglossaries myDoc
\end{verbatim}
(Replace \texttt{myDoc} with the base name of your \LaTeX\ document
file. Avoid spaces in the file name.) If you don't know how to use
the command prompt, you can probably access \texttt{makeglossaries}
via your text editor. Check your editor's manual for advice. If you
are using \texttt{arara} then you can just use the directives:
\begin{verbatim}
% arara: pdflatex
% arara: makeglossaries
% arara: pdflatex
\end{verbatim}

The \texttt{makeglossaries} script is written in Perl, so you need
a~Perl interpreter installed. If you are using a~Unix-like operating
system then you most likely have one installed. If you are using
Windows with the \TeX~Live distribution, then you can use the Perl
interpreter that comes with \TeX~Live. If you are using Windows and
MiK\TeX\ then you need to install a~Perl distribution for Windows.
If you are using \opt3, then you need to do this anyway as
\texttt{xindy} is also written in Perl. If you are using \opt2
and can't work out how to install Perl (or for some reason don't
want to install it) then just use \texttt{makeindex} directly, as
described above.

When sorting the entries, the string comparisons are made according
to each entry's \texttt{sort} key. If this is omitted, the
\texttt{name} key is used. For example, recall the earlier
definition:
\begin{verbatim}
\newglossaryentry{elite}
{
  name={{\'e}lite},
  description={select group or class}
}
\end{verbatim}
No \texttt{sort} key was used, so it's set to the same as the
\texttt{name} key: \verb|{\'e}lite|. How this is interpreted depends
on which option you have used:
\begin{description}
\item[\opt1:] By default, the accent command will be stripped so the
sort value will be \texttt{elite}. This will put the entry in the
\qt{E} letter group. If you use the
\pkgopt[true]{sanitizesort} package option, the sort value will be
interpreted as the sequence of characters: \verb|\| \texttt{'} \texttt{e}
\texttt{l} \texttt{i} \texttt{t} and \texttt{e}. This will place
this entry before the \qt{A} letter group since it starts with a symbol.

\item[\opt2:] The sort key will be interpreted the sequence of characters:
\verb|{| \verb|\| \verb|'| \texttt{e} \verb|}| \texttt{l} \texttt{i} \texttt{t}
and \texttt{e}. The first character is an opening curly brace
\verb|{| so \texttt{makeindex} will put this entry in the ``symbols'' group. 

\item[\opt3:]
\texttt{xindy} disregards \LaTeX\ commands so it sorts on
\texttt{elite}, which puts this entry in the \qt{E} group.
\end{description}


If the \sty{inputenc} package is used and the entry is defined as:
\begin{alltt}
\verb|\newglossaryentry{elite}|
\verb|{|
  name=\verb|{{|\'e\verb|}lite},|
  description=\verb|{select group or class}|
\verb|}|
\end{alltt}
then:
\begin{description}
\item[\opt1:] By default the sort value will be interpreted as
\texttt{elite} so the entry will be put in the \qt{E} letter group.
If you use the \pkgopt[true]{sanitizesort} package option, the
sort value will be interpreted as \texttt{\'elite} where \'e has
been sanitized (so it's no longer an active character) which will
put this entry before the \qt{A} letter group.

\item[\opt2:] \texttt{makeindex} doesn't recognise \texttt{\'e} as
a~letter so it will be put in the symbols group.

\item[\opt3:] \texttt{xindy} will correctly recognise the sort value
\texttt{\'elite} and will place it in whatever letter group is
appropriate for the given language setting. (In English, this would
just be the \qt{E} letter group.)
\end{description}

Therefore if you have extended Latin or non-Latin characters, your
best option is to use \texttt{xindy} (\opt3) with the \sty{inputenc}
package. If you use \texttt{makeindex} (\opt2) you need to specify the 
\texttt{sort} key like this:
\begin{verbatim}
\newglossaryentry{elite}
{
  name={{\'e}lite},
  sort={elite},
  description={select group or class}
}
\end{verbatim}
If you use \opt1, you may or may not need to use the \texttt{sort}
key, but you will need to be careful about fragile commands in the
\texttt{name} key if you don't set the \texttt{sort} key.

\Tableref{tab:optionsp+c} summarises the pros and cons of three options described
above.

\begin{table}[htbp]
 \caption{Glossary Options: Pros and Cons}
 \label{tab:optionsp+c}
 {%
 \centering
 \begin{tabular}{>{\raggedright}p{0.3\textwidth}ccc}
   & \bfseries \opt1 & \bfseries \opt2 & \bfseries \opt3\\
   Requires an external application? &
   \no & \yes & \yes\\
   Requires Perl? &
   \no & \no & \yes\\
   Can sort extended Latin
   or non-Latin alphabets? &
   \no\textsuperscript{\textdagger} & \no & \yes\\
   Efficient sort algorithm? &
   \no & \yes & \yes\\
   Can form ranges in the location lists? &
   \no & \yes & \yes\\
   Can have non-standard locations? &
   \yes & \no & \yes\\
   \ics{newglossaryentry} restricted to preamble? &
   \yes & \no & \no
 \end{tabular}
 \par
 }\textsuperscript{\textdagger} Strips standard \LaTeX\ accents so,
for example, \verb|\AA| is treated the same as A.
\end{table}

\section{Customising the Glossary}
\label{sec:glosstyle}

The default glossary style uses the \texttt{description} environment
to display the entry list. Each entry name is set in the optional
argument of \cs{item} which means that it will typically be
displayed in bold. You can switch to medium weight by redefining
\cs{glsnamefont}:
\begin{verbatim}
\renewcommand*{\glsnamefont}[1]{\textmd{#1}}
\end{verbatim}

By default, a~full stop is appended to the description. To prevent
this from happening use the \texttt{nopostdot} package option:
\begin{verbatim}
\usepackage[nopostdot]{glossaries}
\end{verbatim}

By default, a~location list is displayed for each entry. This refers
to the document locations (for example, the page number) where the
entry has been referenced. If you use \optsor23 described
in \sectionref{sec:printglossaries} location ranges will be compressed.  For
example, if an entry was used on pages~1, 2 and~3, with 
\optsor23 the location list will appear as 1--3, but with \opt1 it
will appear as 1, 2, 3. If you don't want the locations displayed
you can hide them using the \texttt{nonumberlist} package option:
\begin{verbatim}
\usepackage[nonumberlist]{glossaries}
\end{verbatim}

Entries are grouped according to the first letter of
each entry's \texttt{sort} key. By default a~vertical gap is placed
between letter groups. You can suppress this with the
\texttt{nogroupskip} package option:
\begin{verbatim}
\usepackage[nogroupskip]{glossaries}
\end{verbatim}

If the default style doesn't suit your document, you can change the
style using:
\begin{definition}
\cs{setglossarystyle}\marg{style name}
\end{definition}
There are a~number of predefined styles. Glossaries can vary from
a~list of symbols with a~terse description to a~list of words or
phrases with descriptions that span multiple paragraphs, so there's
no ``one style fits all'' solution. You need to choose a~style that
suits your document.

Examples:
\begin{enumerate}
 \item You have entries where the name is a~symbol and the
 description is a~brief phrase or short sentence. Try one of the 
 ``mcol'' styles defined in the \sty{glossary-mcols} package. For example:
\begin{verbatim}
\usepackage[nogroupskip,nopostdot]{glossaries}
\usepackage{glossary-mcols}
\setglossarystyle{mcolindex}
\end{verbatim}

 \item You have entries where the name is a~word or phrase and the
description spans multiple paragraphs. Try one of the ``altlist'' styles. For
example:
\begin{verbatim}
\usepackage[nopostdot]{glossaries}
\setglossarystyle{altlist}
\end{verbatim}

 \item You have entries where the name is a~single word, the
  description is brief, and an associated symbol has been set.
  Use one of the styles that display the symbol (not all of them do). 
  For example, one of the tabular styles:
\begin{verbatim}
\usepackage[nopostdot,nonumberlist]{glossaries}
\setglossarystyle{long4col}
\end{verbatim}
or one of the ``tree'' styles:
\begin{verbatim}
\usepackage[nopostdot,nonumberlist]{glossaries}
\setglossarystyle{tree}
\end{verbatim}
\end{enumerate}

If your glossary consists of a~list of acronyms and you also want to
specify a~description as well as the long form, then you need to use
an acronym style that will suit the glossary style. For example,
use the \texttt{long-short-desc} acronym style:
\begin{verbatim}
\setacronymstyle{long-short-desc}
\end{verbatim}
Define the acronyms with a~description:
\begin{verbatim}
\newacronym
 [description={statistical pattern recognition technique}]
 {svm}{svm}{support vector machine}
\end{verbatim}
Choose a~glossary style that suits wide entry names:
\begin{verbatim}
\setglossarystyle{altlist}
\end{verbatim}

\section{Multiple Glossaries}
\label{sec:multigloss}

The \sty{glossaries} package predefines a~default \texttt{main}
glossary. When you define an entry (using one of the commands
described in \sectionref{sec:defterm}), that entry is automatically
assigned to the default glossary, unless you indicate otherwise 
using the \texttt{type} key. However you first need to
make sure the desired glossary has been defined. This is done using:
\begin{definition}
\cs{newglossary}\oarg{glg}\marg{label}\marg{gls}\marg{glo}\marg{title}
\end{definition}
The \meta{label} is a~label that uniquely identifies this new
glossary. As with other types of identifying labels, be careful not
to use active characters in \meta{label}. The final argument
\meta{title} is the section or chapter heading used by
\cs{printglossary} or \cs{printnoidxglossary}. The other arguments
indicate the file extensions used by
\texttt{makeindex}\slash\texttt{xindy} (described in
\sectionref{sec:printglossaries}). If you use \opt1 described
above, the \meta{glg}, \meta{gls} and \meta{glo} arguments are
ignored. In the case of \optsor23, all glossary definitions must come before
\cs{makeglossaries}.

Since it's quite common for documents to have both a~list of terms
and a~list of acronyms, the \sty{glossaries} package provides the
package option \texttt{acronyms}, which is a~convenient shortcut for
\begin{verbatim}
\newglossary[alg]{acronym}{acr}{acn}{\acronymname}
\end{verbatim}
It also changes the behaviour of \cs{newacronym} so that acronyms
are automatically put in the list of acronyms instead of the main
glossary.

For example, suppose you want a~main glossary for terms, a~list of
acronyms and a~list of notation:
\begin{verbatim}
\usepackage[acronyms]{glossaries}
\newglossary[nlg]{notation}{not}{ntn}{Notation}
\end{verbatim}
After \cs{makeglossaries} (or \cs{makenoidxglossaries}) you can
define the entries. For example:
\begin{verbatim}
\newglossaryentry{gls:set}
{% This entry goes in the `main' glossary
  name=set,
  description={A collection of distinct objects}
}

% This entry goes in the `acronym' glossary:
\newacronym{svm}{svm}{support vector machine}

\newglossaryentry{not:set}
{% This entry goes in the `notation' glossary:
  type=notation,
  name={\ensuremath{\mathcal{S}}},
  description={A set},
  sort={S}}
\end{verbatim}
or if you don't like using \cs{ensuremath}:
\begin{verbatim}
\newglossaryentry{not:set}
{% This entry goes in the `notation' glossary:
  type=notation,
  name={$\mathcal{S}$},
  text={\mathcal{S}},
  description={A set},
  sort={S}}
\end{verbatim}

Each glossary is displayed using:
\begin{definition}
\cs{printnoidxglossary}[type=\meta{type}]
\end{definition}
(\opt1) or
\begin{definition}
\cs{printglossary}[type=\meta{type}]
\end{definition}
(\optsand23). Where \meta{type} is the glossary label. If the
type is omitted the default \texttt{main} glossary is assumed.

There's a~convenient shortcut that will display all the defined
glossaries:
\begin{definition}
\cs{printnoidxglossaries}
\end{definition}
(\opt1) or
\begin{definition}
\cs{printglossaries}
\end{definition}
(\optsand23).

If you use \opt1, you don't need to do anything else. If you use
\optsor23 with the \texttt{makeglossaries} Perl script, you
similarly don't need to do anything else. If you use \optsor23 
without the \texttt{makeglossaries} Perl script then you need to
make sure you run \texttt{makeindex}\slash\texttt{xindy} \emph{for
each defined glossary}. The \meta{gls} and \meta{glo} arguments of
\cs{newglossary} specify the file extensions to use instead of
\texttt{.gls} and \texttt{.glo}. The optional argument \meta{glg} is
the file extension for the transcript file. This should be different
for each glossary in case you need to check for
\texttt{makeindex}\slash\texttt{xindy} errors or warnings if things
go wrong.

For example, suppose you have three glossaries in your document
(\texttt{main}, \texttt{acronym} and \texttt{notation}), 
specified using:
\begin{verbatim}
\usepackage[acronyms]{glossaries}
\newglossary[nlg]{notation}{not}{ntn}{Notation}
\end{verbatim}
Then (assuming your \LaTeX\ document is in a~file called
\texttt{myDoc.tex}):

\begin{description}
\item\opt2:

You need to run \texttt{makeindex} three times:
\begin{verbatim}
makeindex -t myDoc.glg -s myDoc.ist -o myDoc.gls myDoc.glo
makeindex -t myDoc.alg -s myDoc.ist -o myDoc.acr myDoc.acn
makeindex -t myDoc.nlg -s myDoc.ist -o myDoc.not myDoc.ntn
\end{verbatim}

\item\opt3:\nopagebreak

You need to run \texttt{xindy} three times (be careful not to insert
line breaks where the line has wrapped.)
\begin{verbatim}
xindy  -L english -C utf8 -I xindy -M myDoc -t myDoc.glg
-o myDoc.gls myDoc.glo
xindy  -L english -C utf8 -I xindy -M myDoc -t myDoc.alg
-o myDoc.acr myDoc.acn
xindy  -L english -C utf8 -I xindy -M myDoc -t myDoc.nlg
-o myDoc.not myDoc.ntn
\end{verbatim}
\end{description}

\section[glossaries and hyperref]{\styfmt{glossaries} and \styfmt{hyperref}}
\label{sec:hyperref}

Take care if you use the \sty{glossaries} package with
\sty{hyperref}. Contrary to the usual advice that \sty{hyperref}
should be loaded last, \sty{glossaries} must be loaded \emph{after}
\sty{hyperref}:
\begin{verbatim}
\usepackage[colorlinks]{hyperref}
\usepackage{glossaries}
\end{verbatim}
If you use \sty{hyperref} make sure you use PDF\LaTeX\ rather than
the \LaTeX\ to DVI engine. The DVI format can't break hyperlinks across
a~line so long glossary entries (such as the full form of acronyms)
won't line wrap with the DVI engine. Also, hyperlinks in sub- or
superscripts aren't correctly sized with the DVI format.

By default, if the \sty{hyperref} package has been loaded, commands
like \cs{gls} will form a~hyperlink to the relevant entry in the
glossary. If you don't want this to happen for \emph{all} your
glossaries, then use
\begin{definition}
\cs{glsdisablehyper}
\end{definition}
If you want hyperlinks suppressed for entries in specific
glossaries, then use the \texttt{nohypertypes} package option. For
example, if you don't want hyperlinks for entries in the \texttt{acronym} and
\texttt{notation} glossaries but you do want them for entries in the
\texttt{main} glossary, then do:
\begin{verbatim}
\usepackage[colorlinks]{hyperref}
\usepackage[acronym,nohypertypes={acronym,notation}]{glossaries}
\newglossary[nlg]{notation}{not}{ntn}{Notation}
\end{verbatim}

If you want the hyperlinks suppressed the first time an entry is
used, but you want hyperlinks for subsequence references then use
the \texttt{hyperfirst=false} package option:
\begin{verbatim}
\usepackage[colorlinks]{hyperref}
\usepackage[hyperfirst=false]{glossaries}
\end{verbatim}

Take care not to use non-expandable commands in PDF bookmarks. This
isn't specific to the \sty{glossaries} package but is a~limitation
of PDF bookmarks. Non-expandable commands include commands like 
\cs{gls}, \cs{glspl}, \cs{Gls} and \cs{Glspl}. The \sty{hyperref}
package provides a~way of specifying alternative text for the PDF
bookmarks via \cs{texorpdfstring}. For example:
\begin{verbatim}
\section{The \texorpdfstring{\gls{fishage}}{Fish Age}}
\end{verbatim}
However, it's not a~good idea to use commands like \cs{gls} in
a~section heading as you'll end up with the table of contents in 
your location list. Instead you can use
\begin{definition}
\cs{glsentrytext}\marg{label}
\end{definition}
This is expandable provided that the \texttt{text} key doesn't
containing non-expandable code. For example, the following works:
\begin{verbatim}
\section{The \glsentrytext{fishage}}
\end{verbatim}
and it doesn't put the table of contents in the location list.

\section{Cross-References}
\label{sec:xr}

You can add a~reference to another entry in a~location list using
the \texttt{see=}\marg{label} key when you define an entry. 
The referenced entry must also be defined.

For example:
\begin{verbatim}
\longnewglossaryentry{devonian}{name={Devonian}}%
{%
  The geologic period spanning from the end of the 
  Silurian Period to the beginning of the Carboniferous Period.

  This age was known for its remarkable variety of 
  fish species.
}

\newglossaryentry{fishage}
{
  name={Fish Age},
  description={Common name for the Devonian period},
  see={devonian}
}
\end{verbatim}
The cross-reference will appear as \qt{\emph{see} Devonian}. You can
change the \qt{see} tag using the format
\texttt{see=}\oarg{tag}\meta{label}. For example:
\begin{verbatim}
\newglossaryentry{latinalph}
{
  name={Latin alphabet},
  description={alphabet consisting of the letters 
  a, \ldots, z, A, \ldots, Z},
  see=[see also]{exlatinalph}
}
\newglossaryentry{exlatinalph}
{
   name={extended Latin alphabet},
   description={The Latin alphabet extended to include 
   other letters such as ligatures or diacritics.}
}
\end{verbatim}
If you use the \texttt{see} key in the optional argument of
\cs{newacronym}, make sure you enclose the value in braces. For
example:
\begin{verbatim}
\newacronym{ksvm}{ksvm}{kernel support vector machine}
\newacronym
 [see={[see also]{ksvm}}]
 {svm}{svm}{support vector machine}
\end{verbatim}

\section{Further Information}
\label{sec:moreinfo}

Further information can be found in the main \sty{glossaries}
\docref{user manual}{glossaries-user}
and there is also an
\href{http://www.latex-community.org/index.php?option=com_content&view=article&id=263:glossaries-nomenclature-lists-of-symbols-and-acronyms&catid=55:latex-general&Itemid=114}{article on the glossaries package}
on the \urlfootref{http://www.latex-community.org/}{\LaTeX\ Community's}
Know How section and a~chapter on the \sty{glossaries} package in
\href{http://www.dickimaw-books.com/latex/thesis/}{Using
\LaTeX\ to Write a PhD Thesis}.
\end{document}
